### 2019-11-22 0_abstract.tex
Apply Alex's comments.

\begin{abstract}
We present Task Bench, a \emph{parameterized} benchmark designed to
explore the performance of parallel and distributed
programming systems under a variety of application scenarios. Task
Bench lowers the barrier to benchmarking multiple programming systems by making
the implementation for a given system orthogonal to the benchmarks
themselves: every benchmark constructed with Task Bench
runs on every Task Bench implementation. Furthermore,
Task Bench's parameterization enables a wide variety of benchmark
scenarios that distill the key characteristics of larger applications.

>>>>

\begin{abstract}
We present Task Bench, a \emph{parameterized} benchmark designed to
explore the performance of parallel and distributed
programming systems under a variety of application scenarios. Task
Bench lowers the barrier to benchmarking and comparing multiple programming systems by making
the implementation for a given system orthogonal to the benchmarks
themselves: every benchmark constructed with Task Bench
runs on every Task Bench implementation. Furthermore,
Task Bench's parameterization enables a wide variety of benchmark
scenarios that distill the key characteristics of larger applications.

### 2019-11-22 1_introduction.tex
Apply Alex's comments.

Benchmarks in Task Bench are based on the observation that many applications can be modeled
as a set of \emph{tasks}, or coarse-grain units of work, with
dependencies between tasks representing the communication and synchronization required for
distributed execution. A benchmark consists of a \emph{task graph} with tasks for
each point in an \emph{iteration space}, and dependencies
determined by a \emph{dependence relation}. This design permits a
concise description of a wide variety of patterns relevant to
scientific computing and large scale data analysis: trivial parallelism, halo exchanges (such as
seen in structured and unstructured mesh codes), sweeps (such as used
in the discrete ordinates method of radiation simulation), FFTs, trees
(for divide and conquer algorithms), and so on. Tasks execute kernels with a
variety of computational properties, including compute- and
memory-bound operations of varying durations. Dependencies between
tasks can be configured to carry communication payloads of varying size, permitting
the design of communication-bound cases. Finally, multiple
(potentially heterogeneous) task graphs can be executed concurrently
to introduce task parallelism into the workload. Together, these
design elements enable the exploration of a large space of application
behaviors.

Adding a system to Task Bench involves implementing a set of standard
services, such as executing a task or data transfer. Note that, though
the benchmarks are described in terms of task graphs, this is simply a
convenient representation of the computation, and the underlying
system need not provide any native support for tasks. Tasks are simply
coarse units of work; dependencies induce communication. We provide
Task Bench implementations in systems as diverse as MPI and
Spark. Task Bench provides a core API that encapsulates functionality
shared among systems, which reduces implementation effort and makes it
much easier to achieve truly apples-to-apples comparisons between
systems.

This approach has allowed us to incorporate 15 very different parallel
and distributed programming systems (listed in
Table~\ref{tab:systems}).  By running all systems on common benchmarks
we were able to quantify phenomena
that have never before been measured.
Most strikingly, the overheads of the systems we examine vary by more than five orders
of magnitude, with popular, widely used systems at both ends of the spectrum!  Clearly
the slower systems have ``good enough'' performance for some class of applications, while presumably
providing advantages in programmer productivity.

How does one predict whether performance will be good enough for a given application?
Surprisingly, the most commonly reported metrics,
\emph{weak} and \emph{strong} scaling, do not directly characterize
the performance of the underlying
programming system. Weak scaling can hide arbitrary amounts of runtime
system overhead by using sufficiently large problem sizes, and strong
scaling does not separate runtime system overhead from application costs
(such as increased communication) that scale with the number of nodes when
using progressively larger portions of a machine.

>>>>

Benchmarks in Task Bench are based on the observation that regardless
of the programming system in which an application is written, many
applications can be modeled as coarse-grain units of work, called
\emph{tasks}, with dependencies between tasks representing the
communication and synchronization required for parallel and
distributed execution. By explicitly modeling the \emph{task graph}
(with tasks as vertices and dependencies as edges), we make it
possible to explore a wide variety of patterns relevant to
parallel and distributed computing: trivial parallelism, halo exchanges (as
in structured and unstructured mesh codes), sweeps (as
in the discrete ordinates method of simulating radiation), FFTs, trees
(for divide and conquer), DNNs, graph analytics, etc. Tasks execute kernels with a
variety of computational properties, including compute- and
memory-bound loops of varying duration. Dependencies can be configured to carry communication payloads of varying size. Finally, multiple
(potentially heterogeneous) task graphs can be executed concurrently
to introduce task parallelism into the workload. Together, these
elements enable the exploration of a large space of application
behaviors---and make it easy to explore cases limited by runtime
overhead as well as ones where computation or communication is
dominant.

Adding a system to Task Bench involves implementing a set of standard
services, such as executing a task or data transfer. Note that, though
the benchmarks are described in terms of task graphs, this is simply a
convenient representation of the computation, and the underlying
system need not provide any native support for tasks. We provide
Task Bench implementations in systems as diverse as MPI and
Spark. Task Bench provides a core API that encapsulates functionality
shared among systems, which reduces implementation effort and makes it
much easier to achieve truly apples-to-apples comparisons between
systems.

This approach has allowed us to benchmark 15 very different parallel
and distributed programming systems (listed in
Table~\ref{tab:systems}).  By running all systems on common benchmarks
we were able to quantify phenomena
that have never before been measured.
Most strikingly, the overheads of the systems we examine vary by more than five orders
of magnitude, with popular, widely used systems at both ends of the spectrum!  Clearly
the slower systems have ``good enough'' performance for some class of applications, while presumably
providing advantages in programmer productivity.

How does one predict whether performance will be good enough for a given application?
The most commonly reported metrics,
\emph{weak} and \emph{strong} scaling, do not directly characterize
the performance of the underlying
programming system. Weak scaling can hide arbitrary amounts of runtime
system overhead by using sufficiently large problem sizes, and strong
scaling does not separate runtime system overhead from application costs
(such as increased communication) that scale with the number of nodes when
using progressively larger portions of a machine.

### 2019-11-22 2_task_bench.tex
Edits.
Apply Alex's comments.

Task graphs are a combination of an iteration space (with a task for
each point in the space) with a dependence relation.
For simplicity, but without loss of generality, the iteration space in
Task Bench is constrained to be 2-dimensional, with time along
the vertical axis and parallel tasks along the
horizontal. Tasks may depend only on tasks from the immediately
preceding time step. Figure~\ref{fig:task-graphs} shows a number of sample task
graphs that can be implemented with Task Bench. Note that layout is
significant---in particular, column $i$ represents all tasks
with index $i$ in the iteration space over all the time
steps. Generally speaking each column will be
assigned to execute on a different processor core.

Dependencies between tasks are controlled by a configurable dependence
relation. The
dependence relation enumerates the set of tasks from the
previous time step each task depends on. This permits a wide variety
of dependence patterns to be implemented that are relevant to real
applications in high-performance scientific computing and data analysis: stencils,
sweeps, FFTs, trees, etc. Dependence patterns may also be
parameterized, such as picking the $K$ nearest neighbors, or $K$
distant neighbors. Table~\ref{tab:equations} shows equations for the
dependence relations of the patterns in Figure~\ref{fig:task-graphs},
where $t$ is timestep, $i$ is column, and $W$ is the width of the task
graph.

@@

In addition specifying the shape and structure of the task graph, the
Task Bench core API also provides implementations of the kernels
executed by each task as well as other utility routines (to parse
inputs and display results). A sample of the core API compute kernel
is shown in Listing~\ref{lst:compute-kernel}. In addition to reducing
the effort required to implement Task Bench, providing central
implementations of these services ensures that all Task Bench
implementations can be scripted uniformly and eliminates a potential
source of performance disparity that can be a pitfall for mini-apps.

@@

>>>>

Task graphs are a combination of an \emph{iteration space} (with a task for
each point in the space) with a \emph{dependence relation}.
For simplicity, but without loss of generality, the iteration space in
Task Bench is constrained to be 2-dimensional, with time along
the vertical axis and parallel tasks along the
horizontal. Tasks may depend only on tasks from the immediately
preceding time step. Figure~\ref{fig:task-graphs} shows a number of sample task
graphs that can be implemented with Task Bench. Note that layout is
significant: generally speaking each column will be
assigned to execute on a different processor core.

Dependencies between tasks are controlled by a configurable dependence
relation. The
dependence relation enumerates the set of tasks from the
previous time step each task depends on, permitting a wide variety
of patterns to be implemented that are relevant to real
applications: stencils,
sweeps, FFTs, trees, etc. Dependence relations may be
parameterized, such as picking the $K$ nearest neighbors, or $K$
distant neighbors. They may also vary over time, such as in the FFT pattern. The set of dependence relations is extensible, making it easy to add patterns to represent new classes applications. Table~\ref{tab:equations} shows equations for the
dependence relations of the patterns in Figure~\ref{fig:task-graphs},
where $t$ is timestep, $i$ is column, and $W$ is the width of the task
graph.

@@

@@

In addition to specifying the shape and structure of the task graph, the
Task Bench core API also provides implementations of the kernels
executed by each task as well as other utility routines (to parse
inputs and display results). A sample of the core API compute kernel
is shown in Listing~\ref{lst:compute-kernel}. In addition to reducing
the effort required to implement Task Bench, providing central
implementations of these services ensures that all Task Bench
implementations can be scripted uniformly and eliminates a potential
source of performance disparity that can be a pitfall for mini-apps.

### 2019-11-22 4_metg.tex
Apply Alex's comments.

Given that Task Bench permits the rapid exploration of a large space
of application scenarios, one question is how to characterize the
performance and efficiency of the systems under study. As noted
earlier, the overheads of the systems we consider varies by more than
five orders of magnitude. This makes it challenging to extract useful
information from weak and strong scaling runs.

@@

The choice of 50\% is also a parameter and is not fundamental to
METG. We use 50\% in our studies in order to avoid certain pathologies
associated with lower efficiency thresholds (see
Section~\ref{subsec:peak-performance-and-efficiency}), and also
because it aligns with what we observe in practice. For example, in a
CSCS application for new projects, users are instructed to "select the
most parallel efficient job size (ratio of benchmark speed-up
vs. linear speed-up above 50\%)" \cite{CSCSReport}. This corresponds
to METG(50\%).

@@

In Figure~\ref{fig:efficiency-mpi},
efficiency starts at 100\%. Initially task granularity
shrinks with minimal change in efficiency. As tasks shrink further, efficiency drops more rapidly, leading
to a vertical asymptote as overhead dominates useful work.

>>>>

Given that Task Bench permits the rapid exploration of a large space
of application scenarios, one question is how to characterize the
performance and efficiency of the systems under study. As noted
earlier, the overheads of the systems we consider vary by more than
five orders of magnitude, making it challenging to extract useful
information from weak and strong scaling runs.

@@

The choice of 50\% is also a parameter and is not fundamental to
METG. We use 50\% in our studies to avoid pathologies
associated with lower efficiency thresholds (see
Section~\ref{subsec:peak-performance-and-efficiency}), and also
because it aligns with what we observe in practice. For example, in one supercomputer center's application for new projects, users are instructed to "select the
most parallel efficient job size (ratio of benchmark speed-up
vs. linear speed-up above 50\%)" \cite{CSCSReport}. This corresponds
to METG(50\%).

@@

In Figure~\ref{fig:efficiency-mpi},
efficiency starts at 100\%. Initially task granularity
shrinks with minimal change in efficiency. As tasks shrink further, efficiency drops more rapidly, leading
to a vertical asymptote as overhead comes to dominate useful work.

### 2019-11-22 5_evaluation.tex
Fiddle with figure placement.
Move load imbalance back.
Update reference to supplement.
Edit section title.
Apply Alex's comments.

Most systems achieve or nearly achieve peak FLOP/s. Some
systems reserve a number of cores (usually 1 or 2) for internal
runtime usage; these systems take a minor hit in peak FLOP/s compared
to systems which share all cores between the application and runtime. Some of the
higher overhead systems struggle to achieve peak FLOP/s, though in most cases
the curves suggest that performance would continue to improve if we
were to run larger problem sizes. Unfortunately, the excessive
computational cost of running such tests makes this prohibitively
expensive. For example, the Spark job in
this case ran for over 6 hours.

@@

Most systems for HPC are highly scalable,
but this is not true of all the systems included in this
evaluation. Lower is better in Figure~\ref{fig:metg-compute}, and
flat is ideal. Lines that rise with node count indicate less than
ideal scaling. Most notably, Spark is primarily
intended industrial data center applications with task
granularities measured in tens of seconds. Spark uses a centralized
controller, which limits throughput, and this is visible in the figure
as the line for Spark immediately rises with node count. Keep in mind
that Spark is being evaluated here with a non-trivial dependence
pattern that is relatively unrepresentative of Spark's normal use
cases. Spark is more efficient with trivial parallelism, as described
in Section~\ref{subsec:number-of-dependencies}.

@@

\subsection{Communication Hiding}

@@

In order to determine the cost of scheduling tasks on the GPU, Figure~\ref{fig:cuda-efficiency} compares MPI
and MPI+CUDA on the Piz Daint supercomputer. The CUDA
compute kernel achieves a peak performance of $4.759 \times 10^{12}$
FLOP/s, which is very close to the officially reported number
$4.761 \times 10^{12}$. The CPU achieves $5.726 \times 10^{11}$
FLOP/s. Note that the kernels perform different numbers of
operations as the GPU requires more work to reach peak
performance. The x-axis in Figure~\ref{fig:cuda-efficiency} is
normalized to keep FLOPs constant for a given problem
size.

>>>>

Most systems achieve or nearly achieve peak FLOP/s. Some
systems reserve a number of cores (usually 1 or 2) for internal
runtime usage; these systems take a minor hit in peak FLOP/s compared
to systems that share all cores between the application and runtime. Some of the
higher overhead systems struggle to achieve peak FLOP/s, though in most cases
the curves suggest that performance would continue to improve if we
were to run larger problem sizes. Unfortunately, the excessive
computational cost of running such tests makes this prohibitively
expensive. For example, the Spark job in
this case ran for over 6 hours.

@@

Most systems for HPC are highly scalable,
but this is not true of all the systems included in this
evaluation. Lower is better in Figure~\ref{fig:metg-compute}, and
flat is ideal. Lines that rise with node count indicate less than
ideal scaling. Most notably, Spark is primarily
intended for industrial data center applications with task
granularities measured in seconds. Spark uses a centralized
controller, which limits throughput, and this is visible in the figure
as the line for Spark immediately rises with node count. Keep in mind
that Spark is being evaluated here with a non-trivial dependence
pattern that is relatively unrepresentative of Spark's normal use
cases. Spark is more efficient with trivial parallelism, as described
in Section~\ref{subsec:number-of-dependencies}.

@@

\subsection{Overlapping Communication and Computation}

@@

To determine the cost of scheduling tasks on GPUs, Figure~\ref{fig:cuda-efficiency} compares MPI
and MPI+CUDA on the Piz Daint supercomputer. The CUDA
compute kernel achieves a peak performance of $4.759 \times 10^{12}$
FLOP/s, which is very close to the officially reported number
$4.761 \times 10^{12}$. The CPU achieves $5.726 \times 10^{11}$
FLOP/s. Note that the kernels perform different numbers of
operations as the GPU requires more work to reach peak
performance. The x-axis in Figure~\ref{fig:cuda-efficiency} is
normalized to keep FLOPs constant for a given problem
size.

### 2019-11-22 6_related.tex
Edits from Sean.
Fiddle with figure placement.
Apply Alex's comments.

% Force this paragraph to page break at a hyphen because otherwise
% it's going to be nearly impossible to cut enough space out of this
% paper.
\let\oldbrokenpenalty\brokenpenalty
\brokenpenalty=0

@@

Microbenchmarks can be easier to implement, but risk being
unrepresentative of real applications. PRK Stencil~\cite{PRK14}
contains 2 kernels (which combined fit in under 50 lines of C++)
and a halo exchange on a structured grid; an
evaluation~\cite{PRKRuntimes16} compares MPI variants, SHMEM, UPC,
Charm++, and Grappa. The NAS benchmark suite~\cite{NAS91, NAS95}
consist mostly of small kernels for dense array computations and has
implementations in OpenMP~\cite{NASOpenMP99}, MPI and
MPI+OpenMP~\cite{NASMPIOpenMP00}, and
Charm++~\cite{NASCharm96}. Neither PRK nor NAS achieve the breadth of
coverage, flexibility, or ease of implementation of Task Bench. We
believe the evidence for ease of implementation is clear; Task Bench
is implemented for more systems than all previous mini-apps and
microbenchmarks. Task Bench is also not a fixed set of benchmarks, and
it is even easier to use Task Bench to generate new benchmarks.

>>>>

% Force this paragraph to page break at a hyphen because otherwise
% it's going to be nearly impossible to cut enough space out of this
% paper.
%% \let\oldbrokenpenalty\brokenpenalty
\brokenpenalty=0

@@

Microbenchmarks can be easier to implement, but risk being
unrepresentative of real applications. PRK Stencil~\cite{PRK14}
contains 2 kernels (which combined fit in under 50 lines of C++)
and a halo exchange on a structured grid; an
evaluation~\cite{PRKRuntimes16} compares MPI variants, SHMEM, UPC,
Charm++, and Grappa. The NAS benchmark suite~\cite{NAS91, NAS95}
consists mostly of small kernels for dense array computations and has
implementations in OpenMP~\cite{NASOpenMP99}, MPI and
MPI+OpenMP~\cite{NASMPIOpenMP00}, and
Charm++~\cite{NASCharm96}. Neither PRK nor NAS achieve the breadth of
coverage, flexibility, or ease of implementation of Task Bench. We
believe the evidence for ease of implementation is clear; Task Bench
is implemented for more systems than all previous mini-apps and
microbenchmarks. Task Bench is also not a fixed set of benchmarks, and
it is even easier to use Task Bench to generate new benchmarks.
